### 2017-08-22 rnw/Onephase_Estimators and their Application.Rnw
update Andreas 22.08.2017

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{One-phase Estimators and their Application}
\label{sec:onephase_and_appl}

% ---------------------------------------------------------------------------- %
  \subsection{One-phase Estimators}

\subsubsection{Mathematical Background}

\subsubsection{Application}

\newpage