### 2017-10-09 rnw/Analysis and Visualization.Rnw
update Andreas 09.10.2017(spelling errors)

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Analysis and Visualization}
\label{sec:anal_and_vis}

The package \pkg{forestinventory} also provides methods for \textit{analyzing} and \textit{visualizing} the estimation results created by the \code{onephase()}, \code{twophase()} and \code{threephase()} functions.

\subsection{Analysis}

If we want to compare the results and performances of different estimation methods and estimators for a given global or small area inventory, the function to start with is \code{estTable()}. This function restructures the results of \code{onephase()}, \code{twophase()} and \code{threephase()} estimations and merges them into one single data set that provides the basis for further analysis. For demonstration purposes, we will first recalculate the one-phase estimations as well as the two-phase and three-phase \textit{extended pseudo synthetic} and the \textit{pseudo synthetic} estimations for the \code{grisons} data set:

\begin{small}
<<>>=
op <- onephase(formula = tvol~1, data = grisons,
               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
               area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))
@
\end{small}

\begin{small}
<<>>=
extpsynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                        phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                        small_area = list(sa.col = "smallarea", 
                                          areas = c("A", "B","C", "D"),
                                          unbiased = TRUE))
@
\end{small}

\begin{small}
<<>>=
psynth_2p <- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                      phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                      boundary_weights = "boundary_weights",
                      small_area = list(sa.col = "smallarea", 
                                        areas = c("A", "B", "C", "D"),
                                        unbiased = FALSE))
@
\end{small}

\begin{small}
<<>>=
extpsynth_3p <- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area=list(sa.col = "smallarea", areas = c("A", "B", "C", "D"),
                                unbiased = TRUE),
                boundary_weights = "boundary_weights")
@
\end{small}

\begin{small}
<<>>=
psynth_3p <- threephase(formula.s0 = tvol ~ mean,
                formula.s1 = tvol ~ mean + stddev + max + q75, data = grisons,
                phase_id = list(phase.col = "phase_id_3p", s1.id = 1, terrgrid.id = 2),
                small_area = list(sa.col = "smallarea", 
                                  areas = c("A", "B", "C", "D"),
                                  unbiased = FALSE),
                boundary_weights = "boundary_weights")
@
\end{small}

We can then aggregate all estimation objects in a \code{list} and pass it to the \code{estTable()}-function:

\begin{small}
<<>>=
grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                             extpsynth_3p, psynth_3p),
                             sae = TRUE, 
                             vartypes = c("variance", "g_variance",  "ext_variance"))
@
\end{small}

The function merges the estimation results and returns a \code{list} object with the subclasses \code{"esttable"  "smallarea"}. If one prefers the \code{data.frame} format for further analysis, this can easily be done using \code{as.data.frame(grisons.sae.table)}. Note however that the package \pkg{forestinventory} provides several \code{S3} generic methods that particularly respond to the class \code{"esttable"}. Note that the \code{vartypes} argument can be used to restrict the \code{estTable()} output to certain types of variances.

The structure of an \code{esttable} object is very similar to the objects created by the small area estimation functions of the package. However, the point estimates and variances from all estimation objects passed to \code{estTable()} have been stored in one single column (\code{estimate} and \code{variance}) and can be distinguished by the variables \code{method}, \code{estimator} and \code{vartype} which specify the estimation method (one, two or three-phase), the estimator and the type of variance (\code{g_} for \textit{g-weight} and \code{ext_} for \textit{external} variance) that was applied. Per default, also the confidence intervals are added.

\begin{small}
<<eval=FALSE>>=
str(grisons.sae.table)
@
\end{small}

\begin{footnotesize}
<<echo=FALSE>>=
str(grisons.sae.table)
@
\end{footnotesize}

Note that \code{estTable()} also returns the \textbf{\textit{estimation error}} (\code{error}) that is defined as the \textit{standard error} devided by the \textit{point estimate}:

\begin{equation}\label{eq:esterror}
  error_{[\%]}=\frac{\sqrt{\hat{\var}(\hat{Y})}}{\hat{Y}}\cdot100
\end{equation}

As mentioned in the introduction (Section \ref{sec:intro}), the application of multi-phase estimation techniques in forest practice have two main \textbf{goals}:  
\begin{enumerate}
  \item \textit{\textbf{maintaining}} the one-phase estimation precision under reduced costs, i.e. reduced $s_2$ sample size
  \item \textit{\textbf{increasing}} the estimation precision at identical costs, i.e. without increasing the $s_2$ sample size
\end{enumerate}

The function \code{mphase.gain()} was particularly written to quantify the potential benefit of a multi-phase global or small area estimation compared to its respective one-phase estimation. The function takes an \code{esttable} object as input and returns a validation of which multi-phase method and estimator performed best in comparison to the onephase estimation (\textit{baseline}) in terms of estimation precision. If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation. The argument \code{pref.vartype} can be used to define what type of variance (\textit{g-weight} or \textit{external}) should be used for the comparison. Per default (\code{exclude.synth = TRUE}), synthetic estimations (\textit{synth} and \textit{psynth} estimator) are not considered for the comparison since they normally have a much smaller variance but to the cost of a potential bias.

<<echo=FALSE>>=
options(width=10000) 
@

\begin{small}
<<>>=
mphase.gain(grisons.sae.table, pref.vartype = "g_variance")
@
\end{small}

% Interpretation: (gain and rel.eff)
The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multi-phase estimator (\code{var_multiphase}). The multi-phase estimation procedure is again specified in the \code{method} and \code{estimator} column. The last two columns then actually quantify the potential benefit of the multi-phase estimation. The \code{gain} is the reduction (if its value is positive) or possibly also the increase (if its value is negative) in variance when applying the multi-phase as alternative to the one-phase estimation. For small area \code{"B"}, it can be read as "the \textit{twophase extended psynth} estimation procedure leads to a 67.9 \% reduction in variance compared to the one-phase procedure". The column \code{rel.eff} specifies the \textbf{\textit{relative efficiency}} which is defined as the ratio between the one-phase variance and the multi-phase variance:

\begin{equation}\label{eq:releff}
  rel.eff_{[\%]}=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}\cdot100
  \end{equation}

The \textit{relative efficiency} can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator. For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the \textit{twophase extended psynth} estimator. If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple mean of quantifying the financial benefit of multi-phase forest inventories.

\subsection{Visualization}

The package \pkg{forestinventory} also provides a \code{S3} generic plot method based on the \pkg{ggplot2} package \citep{ggplot2} to visualize the estimation results. Likewise \code{mphase.gain()}, also the plot method requires an \code{esttable} object created by the \code{estTable()}-function. The \code{plot()}-function allows for two kinds of visualization: \textbf{1)} the \textit{\textbf{point estimates}} with overlayed confidence intervals, and \textbf{2)} the \textit{\textbf{estimation errors}}.

If a plot of the point estimates with overlayed confidence intervals is required, one has to make sure that the confidence intervals (\code{ci_lower}, \code{ci_upper}) are contained in the \code{esttable} object which is automatically the case using the default setting of \code{estTable()}:

\begin{small}
<<>>=
grisons.sae.table <- estTable(est.list = list(op, extpsynth_2p, psynth_2p,
                                             extpsynth_3p, psynth_3p),
                             sae = TRUE, 
                             vartypes = c("variance", "g_variance",  "ext_variance"))
@
\end{small}

The graphic for the point estimates and confidence intervals is then produced by passing the \code{esstable} object to the \code{S3} generic \code{plot} method and setting the argument \code{yvar = "estimate"}:

\begin{small}
<<eval=FALSE>>=
plot(grisons.sae.table, ncol = 2, yvar = "estimate") +
  ylab("Timber Volume [m3/ha]")
@
\end{small}

\begin{figure}[htb]
\centering
\resizebox{0.9\hsize}{!}{\includegraphics{fig/plot_est_ci.png}}
\end{figure}

If we want the estimation errors to be plotted, we simply set the \code{yvar} argument \code{"error"} (default):

\begin{small}
<<eval=FALSE>>=
plot(grisons.sae.table, ncol = 2)
@
\end{small}

\begin{figure}[htb]
\centering
\resizebox{0.86\hsize}{!}{\includegraphics{fig/plot_error.png}}
\end{figure}

Note that the \code{plot()}-function call can arbitrarily be extended by additional \pkg{ggplot2} parameterizations.

\newpage