### 2017-10-12 rnw/Future Plans.Rnw

% !Rnw root = JStatSoft_forestinventory_master.Rnw

\section{Future Plans}
\label{sec:future}

The \pkg{forestinventory} package currently provides a fairly well-rounded toolkit for forestry inventorists to integrate auxiliary information into their estimates using the model-assisted methods under the design-based approach.  Although 32 combinations of inventory scenarios, estimators and sample designs are covered, there are still potential improvements planned for the future. As this is an open-source project, everyone is encouraged to give feedback and/or make contributions on the package's development page on GitHub \citep{github_forestinventory}. Currently planned extensions include:

\begin{itemize}
\item Implement parallel procedures for efficiently calculating many small areas.
\item Allow functions to accept objects of class \code{data.table} from the \pkg{data.table} package \citep{dt2017} to improve memory efficiency.
\item Enable the user to choose other types of models than linear regressions fitted with OLS.
\end{itemize}