### 2018-09-16 JStatSoft_forestinventory_master-concordance.tex
update andi 16_09_2018(1)
update andi 16_09_2018(1)

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
121 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
127 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
1 2 8 0 1 2 130 1 1 5 7 0 1 2 2 1 1 2 27 0 1 2 3 1 1 2 13 0 1 2 5 1 %
1 5 4 0 1 1 8 0 1 2 7 1 1 5 7 0 1 2 2 1 1 2 9 0 1 2 7 1 1 3 5 0 1 2 %
1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 5 7 0 1 2 2 1 1 2 9 0 %
1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
133}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
102 1 1 3 2 0 1 1 3 0 1 2 4 1 1 5 7 0 1 2 3 1 1 2 12 0 1 2 4 1 1 5 7 %
0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
135 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
1 2 1 1 4 38 1 1 2 11 0 1 2 80 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1309:%
141 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1314:%
1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1549:%
147 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1554:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
0 1 2 2 1 1 6 8 0 1 2 4 1 1 4 6 0 1 2 6 1 1 2 4 0 1 2 40 1 1 2 11 0 %
1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
153 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
159 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
1 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1772:%
165 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1777:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 22 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1956:%
171 11 1}