### 2019-04-27 rnw/Appendix.Rnw

% !Rnw root = JStatSoft_forestinventory_master.Rnw

% set global options for R-output:
<<echo=FALSE>>=
options(prompt = "R> ", continue = "+  ", width = 77, useFancyQuotes = FALSE)
@

\section*{Appendix}
\label{sec:appendix}

\subsection*{Simulation for global two-phase estimation}

<<eval=FALSE>>=
target.suface <- function(x0, y0){
  30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)}
@

<<eval=FALSE>>=
library("rmutil")
int2(target.suface, a = c(0,0), b=c(3,2))/6
int2(target.suface, a = c(0.5, 0.3), b=c(2, 1.3))/(1*1.5) 
@

<<eval=FALSE>>=
sample.generator<- function(n1, n2, target.surface=target.surface){
  realization <- matrix(NA,n1,7)
  for(i in 1:n1){
    x0=2*runif(1)
    y0=3*runif(1)
    realization[i,1] <- target.suface(x0, y0)
    realization[i,2]<-x0
    realization[i,3]<-y0
    realization[i,4]<-x0*x0
    realization[i,5]<-x0*y0
    realization[i,6]<-y0*y0
    realization[i,7]<-0
  }
  realization <- as.data.frame(realization)
  names(realization) <- c("response","x","y","xx","xy","yy", "area")
  realization$phase <- 1
  realization$phase[sample(nrow(realization), n2)] <- 2
  realization$response[realization$phase == 1]<- NA
  realization$area[realization$x >= 0.3 & realization$x <= 1.3 & 
                   realization$y >= 0.5 & realization$y <= 2]<- 1
  return(realization)
}
@

<<eval=FALSE>>=

run.global.sim <- function(num_replications, n1.vec){

  sfrac <- seq(0.02, 0.26, by=0.02) 
  true_value<- 39.17

  coverage_rates <- data.frame(n1=integer(), n2=integer(), 
                               sampling_fractions=numeric(), 
                               avg.error.ext=numeric(),
                               avg.error.g=numeric(),
                               covrate.ext=numeric(),
                               covrate.g=numeric(),
                               no.valid.sims = integer())

  for(i in 1:length(sfrac)){

    n2 <- integer()
    error.ext <- numeric()
    error.g <- numeric()
    ci.ext.logical <- logical()
    ci.g.logical <- logical()
    no.valid.sims. <- integer()

    for(j in 1:num_replications){

      realization <- sample.generator(n1 = n1.vec, n2 = round(n1.vec*sfrac[i]))

      est <- twophase(formula = response ~ y + x + xx, data=realization,
                      phase_id = list(phase.col = "phase", terrgrid.id = 2))

      ci <- confint(est)
      ci.ext.logical <- append(ci.ext.logical, 
                               ci$ci$ci_lower_ext <= true_value & 
                               true_value <= ci$ci$ci_upper_ext)
      ci.g.logical <- append(ci.g.logical, 
                             ci$ci$ci_lower_g <= true_value & 
                             true_value <= ci$ci$ci_upper_g)
      error.ext <- append(error.ext, 
                          sqrt(est$estimation$ext_variance)/est$estimation$estimate)
      error.g <- append(error.g, 
                        sqrt(est$estimation$g_variance)/est$estimation$estimate)

      n2 <- append(n2, est$estimation$n2)
      no.valid.sims. <- sum(n2 > 1)

    }

    coverage_rates <- rbind(coverage_rates, 
                            data.frame(n1=n1.vec, n2=round(mean(n2)), 
                                       sampling_fractions=sfrac[i], 
                                       avg.error.ext=mean(error.ext, na.rm = TRUE),
                                       avg.error.g=mean(error.g, na.rm = TRUE),
                                       covrate.ext=mean(ci.ext.logical, na.rm = TRUE), 
                                       covrate.g=mean(ci.g.logical, na.rm = TRUE),
                                       no.valid.sims=no.valid.sims.))

  }

return(coverage_rates)

} 
@

\newpage

<<eval=FALSE>>=

library(forestinventory)
global.sims<- data.frame()
n1_sims<- c(250, 500, 750, 1000, 1500, 2000) 
  for (i in 1:length(n1_sims)){
    res<- run.global.sim(num_replications = 10000, n1.vec = n1_sims[i])
    global.sims <- rbind(global.sims, res)
  }
@

\newpage